PTFO 8-8695: Two Stars, Two Signals, No Planet

### 2020-06-13 paper_review/response_20200605.txt

Dear Dr. Shafter,

We thank you for organizing the review of our paper, and also
thank the referee for constructive criticism.  We have considered
the referee's remarks carefully and revised our manuscript
accordingly.

@@

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
Overall: It would help if the authors used hours or days
throughout the paper in a consistent manner. The period is
referred to as 0.5d and 11.98h just a couple paragraphs apart
(e.g. in Section 5.1). The abstract uses hours, most plots use
days.

RESPONSE
----------
> Throughout the manuscript, we have shifted the units wherever
> possible to hours.
>
> Exceptions include the ephemeris (which is quoted in days), and
> the periods listed in Figures 3 and 6, which are quoted in
> days. The former is standard convention. The latter was because
> we felt it important to list the periods used to generate the
> figures at a higher precision, and quoting higher precision in
> units of hours seems atypical.

@@

******************************************
REFEREE COMMENT
----------

Introduction: I do not think this is youngest hot Jupiter. That
should be true just for transiting systems, and it was the
youngest at discovery, but the two RV planets from ESPaDOnS are
younger (Donati et al. 2016 and Yu et al. 2017).  Both are hot
Jupiters and in Taurus. I believe the age of Tap 26 is more
ambiguous due to age spreads (multiple groups?) in Taurus, but
the age of V830 Tau is certainly <3 Myr.

RESPONSE
----------

> Thank you for this correction. We have qualified the relevant
> sentence as the youngest known *transiting* hot Jupiter, and
> added a footnote to cite the V830 Tau or TAP 26 results.

@@

******************************************
REFEREE COMMENT
----------

Introduction: I also think it wasn't the first H-alpha detection.
That belongs to LkCa15, which was imaged in H-alpha in 2015
(Sallum et al. 2015). Although that planet might not exist, so
this comment is probably OK. It would still be the first close-in
planet either way. I'd suggest weakening the statement slightly
to err on the side of caution.

RESPONSE
----------
> We followed the referee's suggestion by stating that PTFO 8-8695
> would have been the first close-in planet for which Hα emission
> was detected.

@@

******************************************
REFEREE COMMENT
----------

Introduction: Barnes et al. (2013) claimed that the planet and
star precessed on a period of 300-600 days, and that this could
somewhat fit variable depths and missing transits. The Ciardi et
al. 2015 paper uses this as their model, in addition to gravity
darkening as mentioned in the current introduction. I suggest
mentioning the precession of the planet in the introduction, as
many papers used the precession model or at least discussed it.
The line "...the gravity-darkened star precessed..." seems to
allude to this, but it's unclear from the rest of the text.

RESPONSE
----------
> We have changed the wording of the last sentence of the first
> paragraph in the introduction to use the word
> "precess", rather than "torque".

@@

******************************************
REFEREE COMMENT
----------

Section 2: The first paragraph discusses the full frame images,
while the next section discusses the 2m data. Why did the authors
use the FFIs when they have the (higher quality) 2m data?

RESPONSE
----------
> For the essential analysis of the study (i.e., extracting and
> fitting the long and short signals), we used the 2m SPOC
> lightcurve.
>
> Our discussion of the FFI lightcurves in Section 2 was motivated
> by the discontinuity in the light curve at BJD 2458488.3,
> which was eventually excluded from further analysis. We wanted
> to compare the FFI and 2m light ccurves to see whether this jump
> was an instrumental systematic or a software systematic.
>
> A secondary reason to introduce the concept of FFIs in Section
> 2 is that they are used to measure the variability amplitudes
> of the other members of 25 Ori-1 in Section 4.3.

@@

******************************************
REFEREE COMMENT
----------

Section 2.2.2: "These details concerning the group membership for
one object may seem excessive to those accustomed to the simple
distinction between "young cluster members" and "old field
stars". I suggest removing this. It's not adding anything and
sounds like the authors are making assumptions about the reader.
The membership of this object is essential for an age assessment,
and can/should be updated given the new Gaia data. Since most of
the object class in question are at similar ages, this directly
relates to the study. If it had turned out to be ~100 Myr, we
would all be more skeptical (although ~100 Myr stars like this
are known, they are more rare).

RESPONSE
----------
> We appreciate this suggestion and have omitted this sentence.

@@

******************************************
REFEREE COMMENT
----------

Section 3/4: This deals mostly with identification of the second
star. However, the second star appears to be an assumption of
section 3.2. If the periods come from two stars that could be
separated (e.g., wide binaries), it would have been best to do so
rather than try to model both by including fs and fl terms.  So
it is odd that the binary search is after the light curve model
in the paper. It might make more narrative sense to move the
light curve model to after section 4, since 4 motivates 3.2.

RESPONSE
----------
> We considered re-organizing the paper as the referee suggested,
> but ultimately decided against it.  We chose to present the
> TESS data, and show the 2 signals, which raises the possibility
> of binarity - which is then investigated further with Gaia.
> The referee suggests showing the Gaia evidence for binarity, and then
> showing the TESS data with 2 signals. To us, there does not seem
> to be any advantage to inverting the order.

@@

******************************************
REFEREE COMMENT
----------

Section 4.3: My biggest concern with this analysis is reliance on
the chi^2 from Gaia to assess binarity. This is not the suggested
path from Gaia, as it is subject to a noise model Gaia now knows
has issues. For example, nearly 100% of mid/late M dwarfs have a
high chi^2 (and a large excess astrometric noise), but clearly
they are not all binaries. This is (partially) because of a
color-term in the PSF model that was not initially understood.
Similarly, almost all bright stars have a large RUWE, as do the
bluest stars. The authors correctly use RUWE, the suggested
method, but the RUWE analysis is less convincing and clearly the
authors are leaning on the chi^2.

A good reference on this would be Belokurov et al. 2020, which
shows what kinds of binaries RUWE can and cannot detect. This
binary might simply be too tight to create an astrometric signal
(in which case the RUWE makes sense). The typical cutoff for
binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler et
al. 2019, Pearce et al. 2020, among many others.  I suggest
removing the chi^2 analysis. If the authors want to keep it, they
need to argue why it is a more meaningful metric than RUWE, and
attempt to account for color and other parameters that impact
chi^2.

RESPONSE
----------
> We appreciate this criticism. We've followed the referee's advice
> and replaced this sub-plot with RUWE instead of reduced chi^2.
> The text in Section 4.3 and captions have been updated accordingly.

@@

RESPONSE
----------
> We have updated the text to include this point -- thank you for
> making it.
> 
> We have also updated the text to reflect our hesitation about
> whether or not the phenomenon is associated with binarity,
> primarily because multiple photometric periods do not
> necessarily imply binarity.

******************************************
REFEREE COMMENT
----------

Section 5.4: The authors conclude this subsection with a
reasonable argument that this isn't a protoplanet. Does this same
argument not rule out gas/debris at the Keplerian coronation
radius? What could maintain such clouds for years/decades and
then have them jump phase but at the same period? It's hard to do
this if it is not somehow physically attached to the star in some
what (e.g., spots can do this).

RESPONSE
----------
> This comment caused us to consider more carefully the merits
> and weaknesses of the two hypotheses we discussed. 
> We ended up with a stronger preference for
> clumpy torus model over the spot occultation model. The reasons
> for the preference are given in the text: 1) the spot
> occultation model seems more geometrically contrived than we
> initially thought, and 2) the Halpha excess seen by
> Johns-Krull+16 (and also in unpublished observations by
> Stauffer of scallop shells) seems likely to require hydrogen in
> orbit near the corotation radius.
>
> In response to the question of: "why can the shifting phase be
> explained more easily by dust/gas clumps than the
> protoplanet?", we now refer D'Angelo and Spruit 2012 (see in
> particular their Section 1), which discusses the possibility of
> cyclic accretion, with periods of "stalled" mass accumulation
> outside the corotation radius alternating with accreting
> phases.

### 2020-06-13 paper/ms.tex
Update ms.tex

The explanations that \replaced{are not yet}{have not yet been} ruled
out include {\it (i)} transiting clumps of \added{dust or }gas at the
Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
cold spots on the stellar surface} by an optically thick disk.  The
first and last explanations have added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
possibility of distinct mechanisms explaining these distinct
variability classes remains open.}{ The model of hot and coldspot
occultations by an optically thick disk \citep{zhan_complex_2019} is a
qualitative inverse of the torus of circumstellar material: the
difference is whether to place the ``blobs'' that provide the needed
brightness contrast on the surface of the star, or at the co-rotation
radius.  In the spot occultation model, the observing geometry would
require an inner disk to be projected into a narrow band across the
stellar surface, in order to produce variations with sufficiently
short timescales.  The outer disk would need to either not be present,
or to be misaligned enough, and narrow enough, that it did not fall
along the line of sight. These combined requirements seem
geometrically contrived.  }

\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
the past decade could offer additional hints.  \deleted{Specifically,
}\ptfo's transition between having none, one, and multiple dips per
cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.  It
strains the ``enshrouded protoplanet'' interpretation, because there
are no known processes that cause a planet's orbital phase to jump.
\deleted{The dips would then need to be caused by material that was
somehow disrupted from the planet, but somehow remained co-orbital for
an extended duration. This seems implausible.} \added{ The H$\alpha$
excess observed by \citet{johnskrull_h_2016} also seems like an
important piece of evidence.  The model with circumstellar optically
thick hydrogen gas near the co-rotation radius readily explains this
excess.  }

>>>>

The explanations that \replaced{are not yet}{have not yet been} ruled
out include {\it (i)} transiting clumps of \added{dust or }gas at the
Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
cold spots on the stellar surface} by an optically thick disk.  The
first and last explanations have added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
possibility of distinct mechanisms explaining these distinct
variability classes remains open.}{ In the spot occultation scenario
\citep{zhan_complex_2019},
the ``blobs'' that provide the necessary
brightness contrast are on the surface of the star, rather than at the co-rotation
radius. The observing geometry would
require an inner ring to be projected into a narrow band in front of the
stellar disk, in order to produce variations with sufficiently
short timescales. 
% The outer disk would need to either not be present,
% or to be misaligned enough, and narrow enough, that it did not fall
% along the line of sight.
% JNW Doesn't this same problem affect the "dusty torus" model?
This geometric requirement seems too contrived to occur in even
$\lesssim$1\% of young M dwarf systems. }

\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
the past decade could offer additional hints.  \deleted{Specifically,
}\ptfo's transition between having none, one, and multiple dips per
cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
%It
%strains the ``enshrouded protoplanet'' interpretation, because there
%are no known processes that cause a planet's orbital phase to jump.
% JNW Doesn't seem worth beating on a dead horse.
\deleted{The dips would then need to be caused by material that was
somehow disrupted from the planet, but somehow remained co-orbital for
an extended duration. This seems implausible.} \added{ The H$\alpha$
excess observed by \citet{johnskrull_h_2016} also seems like an
important piece of evidence.  The model with circumstellar optically
thick hydrogen gas near the co-rotation radius readily explains this
excess.  }
